---
title: The R version of "introduction_to_python.ipynb"
output: html_document
---

# Introduction to ~~Python~~ R

**Q1.** Write some code which multiplies $12$ and $13$ and assign the result to variable called `result`.

```{r}

result <- 12 * 13
result

```

**Q2.** Two integers, `a` and `b`, are provided in the following cell - run this cell and do not change the values of `a` and `b`. Next, write some code in the subsequent cell that divides `a` by `b` and assigns the result to a variable called `division`.

```{r}
a <- 17
b <- 3
```

```{r}
division = a / b
division
```


**Q3.** Now we want the result as an integer (rounded down). We can do this with an operation known as "floor division".

Write code in the cell below which divides `a` by `b` using "floor division" and assign the result to a variable called `floor`.

```{r}
floor = a %/% b
floor
```

**Q4.** Finally, write code which returns the remainder of dividing `a` by `b` and assign this to a variable called `remainder`.

```{r}
# Add your code below
# remainder = ...
remainder = a %% b
remainder
```
